---
title: "Notes"
author: "Frank Jung"
date: "18 September 2015"
output: 
  html_document: 
    highlight: tango
    theme: united
    toc: yes
---

# Context

You work for Motor Trend, a magazine about the automobile industry. Looking at a
data set of a collection of cars, they are interested in exploring the
relationship between a set of variables and miles per gallon (MPG) (outcome).
They are particularly interested in the following two questions:

* Is an automatic or manual transmission better for MPG?
* Quantify the MPG difference between automatic and manual transmissions?

# Question

Take the mtcars data set and write up an analysis to answer their question using
regression models and exploratory data analyses.

Your report must be:

* Written as a PDF printout of a compiled (using knitr) R markdown document.
* Brief. Roughly the equivalent of 2 pages or less for the main text. Supporting figures in an appendix can be included up to 5 total pages including the 2 for the main report. The appendix can only include figures.
* Include a first paragraph executive summary.

Upload your PDF by clicking the Upload button below the text box.

# Peer Grading

* The criteria that your classmates will use to evaluate and grade your work are shown below. 
* Each criteria is binary: (1 point = criteria met acceptably; 0 points = criteria not met acceptably)
* Your Course Project score will be the sum of the points and will count as 40% of your final grade in the course. 

# Rubric

* Did the student interpret the coefficients correctly?
* Did the student do some exploratory data analyses?
* Did the student fit multiple models and detail their strategy for model selection?
* Did the student answer the questions of interest or detail why the question(s) is (are) not answerable?
* Did the student do a residual plot and some diagnostics?
* Did the student quantify the uncertainty in their conclusions and/or perform an inference correctly?
* Was the report brief (about 2 pages long) for the main body of the report and no longer than 5 with supporting appendix of figures?
* Did the report include an executive summary?
* Was the report done in Rmd (knitr)?

# Workflow

## Executive Summary

* what conclusions can be drawn?
* is a/m a good predictor for performance of mpg?

## Explorartory Anaylsis

* id variables as categorical, ordinal, quantitative
* are there important outliers?
* test for independence of data (confounding variables?)
* select orthogonal variables
* confirm selection meets assumption for fit to a linear model

Good regressors have three main qualities:

* They vary enough along the x axis.
* When they vary along the x axis, the outcome variable varies along the y axis in linear or curvilinear or some identifiable pattern.
* Using the linear pattern as an example, the points stay fairly close to that line.

## Choose Model

* perform multi-co-linearity tests to identify significant regressors

## Model

* summarise model with confidence interval
* provide model coefficients for prediction
* estimate adjusted r-squared to penalize big models (can we get to 70%?)
* residuals
* confidence intervals for coefficients
* confidence intervals for predictors

# References

Choosing a model by Whattest Stats (YouTube)

* [Exploratory Analysis(Multiple linear regression in R)](https://youtu.be/bfGhfalZR-Y)

* [Building a model(Multiple linear regression in R)](https://youtu.be/itEyHjNjMIc)

* [Henderson and Velleman (1981), Building multiple regression models interactively. Biometrics, 37, 391–411.](https://github.com/frankhjung/regmods-project/blob/master/mtcars-henderson.pdf)

For assumptions and good overall approach see 

* [Linear Regression (Kircher)](file:///home/frank/study/regmods-032/docs/linear-regression-kirchner.pdf)

For more detailed treatment of all topics including variable selection see

* [Practical regression and ANOVA](file:///home/frank/study/regmods-032/docs/practical-regression-and-anova-faraway.pdf)

For tests to select best regressors:
* [Multiple (Linear) Regression](http://www.statmethods.net/stats/regression.html)

* [Selecting the best model for multiple linear regression](file:///home/frank/study/regmods-032/docs/linear-regression-model-selection.pdf)

* [MASS step AIC](https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/stepAIC.html)


# To Read

* https://beckmw.wordpress.com/2013/02/05/collinearity-and-stepwise-vif-selection/
* http://stats.stackexchange.com/questions/58141/interpreting-plot-lm
* http://blog.minitab.com/blog/adventures-in-statistics/how-to-interpret-a-regression-model-with-low-r-squared-and-low-p-values
* http://www.statmethods.net/stats/rdiagnostics.html

# Code tests

```{r setoptions, echo=FALSE}
require(knitr, quietly = TRUE)
require(ggplot2, quietly = TRUE)
require(scales, quietly = TRUE)

knitr::opts_chunk$set(echo = TRUE, cache = TRUE, cache.path = "cache/", 
                      fig.path = "figure/", fig.align = "center")

# load cars dataset
require(datasets, quietly = TRUE)
data(mtcars)
```

## exploratory analysis

```{r summary}
str(mtcars)
kable(head(mtcars),align = 'c')
```

## lm fit against all

```{r fitall}
fit <- lm(mpg ~ ., mtcars)
summary(fit)
par(mfrow = c(2, 2))
plot(fit)
```

## correlation

```{r correlation}
round(cor(mtcars), 3)
```

## fitted vs residual

```{r fitbest}
fit <- lm(mpg ~ wt + qsec + am, mtcars)
summary(fit)
round(cor(mtcars[, c("mpg", "wt", "qsec", "am")]), 3)
qplot(fitted(fit), resid(fit))
```

## boxplot

```{r boxplot}
boxplot(mpg ~ am, mtcars, xaxt = "n", xlab = "Transmission Type", ylab = "Miles/(US) gallon")
with(mtcars, points(am + 1, mpg, pch = 19, col = ((am == 0)*1 + 1)))
axis(1, at = c(1, 2), labels = c("automatic", "manual"))
title(main = "Milage by Transmission Type")
```

## ggpairs

```{r ggpairs}
# require(GGally)
# require(ggplot2)
# ggpairs(mtcars, lower = list(continuous = "smooth"), params = c(method = "loess"))
```

## lm plots

```{r statsfit1}
round(cor(mtcars[, c("mpg", "am", "wt", "qsec")]), 3)
fitsum <- summary(fit)
par(mfrow = c(2, 2))
plot(fit)
round(fitsum$adj.r.squared, 3)
```

# What is AIC?

AIC is a goodness of fit measure that favours smaller residual error in the 
model, but penalises for including further predictors and helps avoiding 
overfitting. When comparing models fitted by maximum likelihood to the same
data, the smaller the AIC, the better the fit.

```{r aic}
AIC(fit)
```

## stepwise with AIC

```{r stepwise}
# stepwise regression
library(MASS)
fit <- lm(mpg ~ ., mtcars)
step <- stepAIC(fit, direction = "both")
# display results
step$anova 
```

# Confidence Interval

Now let’s do the slope:

```r
> (sumCoef[2,1] + c(-1, 1) * qt(.975, df = fit$df) * sumCoef[2, 2]) / 10
[1] 355.6 388.6
```
## null hypothesis and t-test

```{r ttestmeans, echo=FALSE}
h0 <- t.test(mpg ~ am, data = mtcars)
```

First lets test if there is a significant difference in mean milage between 
automatic and manual transmission types. Using a Welch two sample t-test for a 
null hypothesis ($H_0$) that the mean milage is the same for automatic and
manual, returns a significant p-value of about `r round(h0$p.value, 4)`.

# Compare Models

## simple vs multi linear regression

```{r}
require(knitr, quietly = TRUE)
require(datasets, quietly = TRUE)
data(mtcars)

modelSLR <- lm(mpg ~ am, mtcars)
modelSLRsum <- summary(modelSLR)

modelMLR <- lm(mpg ~ wt + qsec + am, mtcars)
modelMLRsum <- summary(modelMLR)

fit.models <- c("Simple Linear Regression", "Multi Linear Regression")
fit.r.squared <- c(modelSLRsum$adj.r.squared, modelMLRsum$adj.r.squared)
fit.vif <- 1/(1 - fit.r.squared)
fit.var.residuals <- c(modelSLRsum$sigma, modelMLRsum$sigma)
fit.aic <- c(AIC(modelSLR), AIC(modelMLR))
fit.mean.residuals <- c(mean(modelSLR$residuals), mean(modelMLR$residuals))

# fit.cov <- c(cov(modelSLR$residuals, mtcars$mpg), cov(modelMLR$residuals, mtcars$mpg))

df <- data.frame("Model" = fit.models,
                 "R Squared" = fit.r.squared,
                 VIF = fit.vif,
                 "Residual Variance" = fit.var.residuals,
                 AIC = fit.aic)
kable(df, caption = "Compare Models")

kable(anova(modelSLR, modelMLR), caption = "ANOVA")
```

P-value for 2nd model is significant further supporting the multi linear
regression model.

```{r}
mean(modelSLR$residuals)
mean(modelMLR$residuals)

modelSLRsum$cov.unscaled
modelMLRsum$cov.unscaled
```
