---
title: "Notes"
author: "Frank Jung"
date: "18 September 2015"
output: 
  html_document: 
    highlight: tango
    theme: united
    toc: yes
---

# Context

You work for Motor Trend, a magazine about the automobile industry. Looking at a
data set of a collection of cars, they are interested in exploring the
relationship between a set of variables and miles per gallon (MPG) (outcome).
They are particularly interested in the following two questions:

* Is an automatic or manual transmission better for MPG?
* Quantify the MPG difference between automatic and manual transmissions?

# Question

Take the mtcars data set and write up an analysis to answer their question using
regression models and exploratory data analyses.

Your report must be:

* Written as a PDF printout of a compiled (using knitr) R markdown document.
* Brief. Roughly the equivalent of 2 pages or less for the main text. Supporting figures in an appendix can be included up to 5 total pages including the 2 for the main report. The appendix can only include figures.
* Include a first paragraph executive summary.

Upload your PDF by clicking the Upload button below the text box.

# Peer Grading

* The criteria that your classmates will use to evaluate and grade your work are shown below. 
* Each criteria is binary: (1 point = criteria met acceptably; 0 points = criteria not met acceptably)
* Your Course Project score will be the sum of the points and will count as 40% of your final grade in the course. 

# Rubric

* Did the student interpret the coefficients correctly?
* Did the student do some exploratory data analyses?
* Did the student fit multiple models and detail their strategy for model selection?
* Did the student answer the questions of interest or detail why the question(s) is (are) not answerable?
* Did the student do a residual plot and some diagnostics?
* Did the student quantify the uncertainty in their conclusions and/or perform an inference correctly?
* Was the report brief (about 2 pages long) for the main body of the report and no longer than 5 with supporting appendix of figures?
* Did the report include an executive summary?
* Was the report done in Rmd (knitr)?

# Workflow

## Explorartory Anaylsis

* id variables as categorical, ordinal, quantitative
* are there important outliers?
* test for independence of data (confounding variables?)
* select orthogonal variables
* confirm selection meets assumption for fit to a linear model

Good regressors have three main qualities:

* They vary enough along the x axis.
* When they vary along the x axis, the outcome variable varies along the y axis in linear or curvilinear or some identifiable pattern.
* Using the linear pattern as an example, the points stay fairly close to that line.

## Choose Model

* perform multi-co-linearity tests to identify significant regressors

## Model

* summarise model with confidence interval
* provide model coefficients for prediction
* estimate adjusted r-squared to penalize big models (can we get to 70%?)
* residuals
* confidence intervals for coefficients
* confidence intervals for predictors

## Executive Summary

* what conclusions can be drawn?
* is a/m a good predictor for performance of mpg?

# References

Choosing a model by Whattest Stats (YouTube)

* [Exploratory Analysis(Multiple linear regression in R)](https://youtu.be/bfGhfalZR-Y)
* [Building a model(Multiple linear regression in R)](https://youtu.be/itEyHjNjMIc)

For assumptions and good overall approach see 

* [Linear Regression (Kircher)](file:///home/frank/study/regmods-032/docs/linear-regression-kirchner.pdf)

For more detailed treatment of all topics including variable selection see

* [Practical regression and ANOVA](file:///home/frank/study/regmods-032/docs/practical-regression-and-anova-faraway.pdf)

Model selection

For tests to select best regressors:
* [Multiple (Linear) Regression](http://www.statmethods.net/stats/regression.html)

* [Selecting the best model for multiple linear regression](file:///home/frank/study/regmods-032/docs/linear-regression-model-selection.pdf)

# Code tests

```{r summary}
library(datasets)
data("mtcars")
str(mtcars)
```

```{r fitall}
fit <- lm(mpg ~ ., mtcars)
summary(fit)
par(mfrow = c(2, 2))
plot(fit)
```

```{r correlation}
round(cor(mtcars), 3)
```

```{r stepwise}
# stepwise regression
library(MASS)
fit <- lm(mpg ~ ., mtcars)
step <- stepAIC(fit, direction = "both")
# display results
step$anova 
```

```{r fitbest}
fit <- lm(mpg ~ wt + qsec + am, mtcars)
summary(fit)
round(cor(mtcars[, c("mpg", "wt", "qsec", "am")]), 3)
par(mfrow = c(2, 2))
plot(fit)
```