---
title: 'Motor Trend: Which is better Manual or Automatic?'
author: "Frank Jung"
date: "21 September 2015"
output: 
  pdf_document: 
    fig_caption: yes
    fig_crop: no
    latex_engine: xelatex
    toc: yes
classoption: legalpaper    
---

# Executive Summary

The Executive Summary is a short section that summarizes the purpose, goals, and
approach of your design project. Reading this summary should give any evaluator
(or your boss) a clear idea of the problem you are tackling, your approach to
solving it and roughly where you are in the process. Consider it an abstract or
overview of your project.

Context: for Motor Trend, a magazine about the automobile industry. Looking at a
data set of a collection of cars, they are interested in exploring the 
relationship between a set of variables and miles per gallon (MPG) (outcome). 
They are particularly interested in the following two questions:

* Is an automatic or manual transmission better for MPG?
* Quantify the MPG difference between automatic and manual transmissions?

# Exploratory Analysis

```{r setoptions, echo=FALSE, message=FALSE, warning=FALSE}
require(knitr, quietly = TRUE)
require(ggplot2, quietly = TRUE)
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, cache.path = "cache/", fig.path = "figure/", fig.align = "center")
```
```{r initial, echo=FALSE, message=FALSE, warning=FALSE}
require(datasets)
data(mtcars)
attach(mtcars)
```

The cars dataset contains three categorical observations 


# Choose Model


# References

* [GitHub repository containing project code](https://github.com/frankjung/regmods-project1)
* Henderson and Velleman (1981), Building multiple regression models interactively. Biometrics, 37, 391–411.
* [MT Cars Dataset](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html)
* V/S - [V-shaped](https://en.wikipedia.org/wiki/V_engine) / [Straight](https://en.wikipedia.org/wiki/Straight_engine) engines

# Appendices

## Exploratory Analysis

```{r explore, echo=TRUE, message=FALSE, fig.cap="Exploratory Analysis"}
str(mtcars)

summary(mtcars)

summary(lm(mpg ~ . - 1, mtcars))$coef

boxplot(mpg ~ am, mtcars, xaxt = "n", xlab = "Transmission Type", ylab = "Miles/(US) gallon")
points(am + 1, mpg, pch = 19, col = ((am == 0)*1 + 1))
axis(1, at = c(1, 2), labels = c("automatic", "manual"))
title(main = "Milage by Transmission Type")
```


## Choose Model

```{r stepwise, fig.cap="Choose Model"}
# stepwise regression
library(MASS)
fit <- lm(mpg ~ ., mtcars)
step <- stepAIC(fit, direction = "both")
# display results
step$anova 
```

```{r bestfit, fig.cap="Best Model Fit"}
fit <- lm(mpg ~ wt + qsec + am, mtcars)
summary(fit)
round(cor(mtcars[, c("mpg", "wt", "qsec", "am")]), 3)
par(mfrow = c(2, 2))
plot(fit)
```
