---
title: "Motor Trend: Which is better Manual or Automatic?"
# author: "Frank Jung"
# date: "21 September 2015"
output:
  pdf_document:
    fig_caption: no
    fig_crop: no
    latex_engine: xelatex
#    toc: yes
classoption: legalpaper
---

# Executive Summary

This report seeks to answer the question: Which is better for car milage: 
automatic or manual? Furthermore, can we quantify any milage differences between
these two transmission types? To answer this question _Motor Trend_ has surveyed
various performance metrics. In this report we will show that of the many 
automobile attributes surveyed by _Motor Trend_, the inclusion of just a few 
namely, Transmission Type, Weight and $\frac {1} {4}$ mile time, provide a 
better prediction of milage than using just the transmission type.

```{r setoptions, echo=FALSE}
require(knitr, quietly = TRUE)
require(ggplot2, quietly = TRUE)
require(scales, quietly = TRUE)
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, cache.path = "cache/", 
                      fig.path = "figure/", fig.align = "center")
# load cars dataset
require(datasets, quietly = TRUE)
data(mtcars)
```

# Exploratory Analysis

A data frame with 32 observations on 11 variables.

| Field | Description               |
|-------|---------------------------|
| mpg   | Miles/(US) gallon         |
| cyl   | Number of cylinders (`r sort(unique(mtcars$cyl))`) |
| disp  | Displacement (cu.in.)     |
| hp    | Gross horsepower          |
| drat  | Rear axle ratio           |
| wt    | Weight (lb/1000)          |
| qsec  | 1/4 mile time             |
| vs    | [V shaped engine](https://en.wikipedia.org/wiki/V_engine), [Straight engine](https://en.wikipedia.org/wiki/Straight_engine) |
| am    | Transmission (0 = automatic, 1 = manual) |
| gear  | Number of forward gears (`r sort(unique(mtcars$gear))`) |
| carb  | Number of carburetors (`r sort(unique(mtcars$carb))`) |

The sample of the first rows from dataset looks like:

```{r head}
kable(head(mtcars), caption = "Sample from mtcars")
```

How does milage compare to transmission? The [Milage by Transmission
Type](#appendices) plot would indicate that manual transmissions have better
milage. However, this could be due to other confounding factors. We will
investigate this further in the following sections.

# Simple Linear Model

Consider a simple linear model where we predict the milage (`mpg`) using just 
the categorical variable, transmission (`am`). This returns the coefficients

```{r model1}
model1 <- lm(mpg ~ am, mtcars)
model1sum <- summary(model1)
kable(model1sum$coef, caption = "Model Coefficients")
```

Reassuringly the P-value of the slope coefficient is
`r prettyNum(round(model1sum$coef[2,4], 5))`, which indicates there is some 
amount of linear relationship between milage and transmission type. However, the
predictions made by this model are just the milage sample mean for the 
respective transmission types:

| Transmission | Prediction | Milage Sample Mean |
|--------------|------------|--------------------|
| Automatic | `r round(predict(model1, data.frame(am = 0)), 3)` | `r round(mean(with(mtcars, mtcars[am == 0, "mpg"])), 3)` | 
| Manual | `r round(predict(model1, data.frame(am = 1)), 3)` | `r round(mean(with(mtcars, mtcars[am == 1, "mpg"])), 3)` |

Furthermore, only `r round(model1sum$adj.r.squared * 100, 3)`% of the variation 
($R^2$) is explained by this model. Can we do better? 

# Multivariable Linear Model

As a alternative look at a multi-variable model. To assist us we will use the
AIC step-wise model selection from the MASS library. AIC is a goodness of fit
measure that favours smaller residual error in the model, but penalises for
including further predictors and helps avoiding overfitting. When comparing
models fitted by maximum likelihood to the same data, the smaller the AIC, the
better the fit. The previous simple linear model had a maxium likelihood
(residula standard error) of `r prettyNum(round(model1sum$sigma, 3))` and an AIC
of `r round(AIC(model1), 2)`.

We will perform a stepwise regression using AIC.

```{r stepwise}
library(MASS)
fit <- lm(mpg ~ ., mtcars)
step <- stepAIC(fit, direction = "both", trace = 0)
# step$anova
```

The best model found is: $mpg \sim wt + qsec + am$. The table below summarises
its coefficients:

```{r model2}
model2 <- lm(mpg ~ wt + qsec + am, mtcars)
model2sum <- summary(model2)
kable(model2sum$coef, caption = "Model Coefficients")
```

How does this model compare to our [Simple Linear Model](#simple-linear-model)?
Below is a table to some of the key statistics we can use to compare these two
models.

```{r compare}
fit.models <- c("model1", "model2")
fit.r.squared <- c(model1sum$adj.r.squared, model2sum$adj.r.squared)
fit.var.residuals <- c(model1sum$sigma, model2sum$sigma)
fit.aic <- c(AIC(model1), AIC(model2))
df <- data.frame("Model" = fit.models, "R-Squared" = fit.r.squared, "Residual Variance" = fit.var.residuals, AIC = fit.aic)
kable(df)
```

# References

* [GitHub repository containing project code](https://github.com/frankjung/regmods-project1)
* [Henderson and Velleman (1981), Building multiple regression models interactively. Biometrics, 37, 391–411.](https://github.com/frankhjung/regmods-project/blob/master/mtcars-henderson.pdf)
* [MT Cars Dataset](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html)

# Appendices

```{r violinplot, echo = TRUE, fig.cap = "Milage by Transmission Type"}
mtcars2 <- data.frame(mtcars)
mtcars2$am <- factor(mtcars2$am, labels = c("automatic", "manual"))
ggplot(mtcars2, aes(x = am, y = mpg)) +
    geom_violin() +
    geom_boxplot(width = 0.2, fill = "purple", alpha = 0.4) +
    stat_summary(fun.y = mean, geom = "point", colour = "white", shape = 19, size = 2) +
    labs(x = "Transmission Type", y = "Miles/(US) gallon")
```

```{r redismodel1, fig.cap = "Fitted vs Residuals for model 2", message = FALSE, warning = FALSE}
#plot(fitted(model1), resid(model1))
qplot(fitted(model1), resid(model1)) +
    geom_hline(yintercept = 0, colour = "red") +
    geom_point(shape = 19, size = 2, colour = "purple") +
    xlab("Fitted values") +
    ylab("Residuals") +
    ggtitle("Model 1 fitted values vs residuals")
```

```{r redismodel2, fig.cap = "Fitted vs Residuals for model 2", message = FALSE, warning = FALSE}
# plot(fitted(model2), resid(model2))
qplot(fitted(model2), resid(model2)) +
    geom_hline(yintercept = 0, colour = "red") +
    geom_point(shape = 19, size = 2, colour = "purple") +
    xlab("Fitted values") +
    ylab("Residuals") +
    ggtitle("Model 2 fitted values vs residuals")
```
