---
title: 'Motor Trend: Which is better Manual or Automatic?'
author: "Frank Jung"
date: "21 September 2015"
output:
  pdf_document:
    fig_caption: yes
    fig_crop: no
    latex_engine: xelatex
    toc: yes
classoption: legalpaper
---

# Executive Summary

This report seeks to answer the question: Which is better for car milage:
automatic or manual? Furthermore, can we quantify any milage differences between
these two transmission types? To answer this question _Motor Trend_ has surveyed
various performance metrics. In this report we will show that of the many automobile
attributes surveyed by _Motor Trend_, the inclusion of just a few namely,
Transmission Type, Weight and 1/4 mile time, provide a better prediction of
milage than using just the transmission type.

```{r setoptions, echo=FALSE}
require(knitr, quietly = TRUE)
require(ggplot2, quietly = TRUE)
require(scales, quietly = TRUE)
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, cache.path = "cache/", fig.path = "figure/", fig.align = "center")

require(datasets, quietly = TRUE)
data(mtcars)
```

# Exploratory Analysis

A data frame with 32 observations on 11 variables.

| Field | Description               |
|-------|---------------------------|
| mpg   | Miles/(US) gallon         |
| cyl   | Number of cylinders (`r sort(unique(mtcars$cyl))`) |
| disp  | Displacement (cu.in.)     |
| hp    | Gross horsepower          |
| drat  | Rear axle ratio           |
| wt    | Weight (lb/1000)          |
| qsec  | 1/4 mile time             |
| vs    | [V shaped engine](https://en.wikipedia.org/wiki/V_engine),  [Straight engine](https://en.wikipedia.org/wiki/Straight_engine) |
| am    | Transmission (0 = automatic, 1 = manual) |
| gear  | Number of forward gears (`r sort(unique(mtcars$gear))`) |
| carb  | Number of carburetors (`r sort(unique(mtcars$carb))`) |

The sample of the first rows from dataset looks like:

```{r head, fig.cap="Exploratory Analysis - Sample"}
kable(head(mtcars))
```

How does milage compare to transmission? The boxplot below would indicate that manual transmissions have better milage. However, this could be due to other confounding factors. We will investigate this further in the next section. 

```{r boxplot, fig.cap="Milage by Transmission Type"}
boxplot(mpg ~ am, mtcars, xaxt = "n", xlab = "Transmission Type", ylab = "Miles/(US) gallon")
with(mtcars, points(am + 1, mpg, pch = 19, col = ((am == 0)*1 + 1)))
axis(1, at = c(1, 2), labels = c("automatic", "manual"))
title(main = "Milage by Transmission Type")
```

# Simple Linear Model

Consider a simple linear model where we predict the milage (`mpg`) using just
the categorical variable, transmission (`am`). This returns the coefficients

```{r fit0coef}
fit0 <- lm(mpg ~ am, mtcars)
fit0sum <- summary(fit0)
kable(fit0sum$coef)
```

Reassuringly the P-value of the slope coefficient is `r fit0sum$coef[2,4]` 
which indicates there is some amount of linear relationship between milage and
transmission type. However, the predictions made by this model are just the
milage sample mean for the respective transmission types:

| Transmission | Prediction | Milage Sample Mean |
|--------------|------------|--------------------|
| Automatic | `r round(predict(fit0, data.frame(am = 0)), 3)` | `r round(mean(with(mtcars, mtcars[am == 0, "mpg"])), 3)` | 
| Manual | `r round(predict(fit0, data.frame(am = 1)), 3)` | `r round(mean(with(mtcars, mtcars[am == 1, "mpg"])), 3)` | 

Furthermore, only `r round(fit0sum$adj.r.squared * 100, 3)`% of the variation 
($R^2$) is explained by this model. Can we do better?

# Multivariable Linear Model

As a alternative look at a multi-variable model. 

```{r stepwise}
# stepwise regression to determine best model given all observations
library(MASS)
fit <- lm(mpg ~ ., mtcars)
step <- stepAIC(fit, direction = "both", trace = 0)
# display results
step$anova
```

Better Fit

```{r betterfit}
# better model than mpg ~ am
fit1 <- lm(mpg ~ am + wt + qsec, mtcars)
fit1summary <- summary(fit)
round(fit1summary$r.squared, 3)
round(cor(mtcars[, c("mpg", "am", "wt", "qsec")]), 3)
par(mfrow = c(2, 2))
plot(fit)
```

# References

* [GitHub repository containing project code](https://github.com/frankjung/regmods-project1)
* [Henderson and Velleman (1981), Building multiple regression models interactively. Biometrics, 37, 391–411.](https://github.com/frankhjung/regmods-project/blob/master/mtcars-henderson.pdf)
* [MT Cars Dataset](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html)
* [Report session information](#session_information), Figure \ref{fig:si}

# Appendices

## Session Information

```{r session}
sessionInfo()
```